---
title: "Assignment3"
author: "Liam McCallum"
date: "3/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##1.1
a. The Sample space for the experiment is 16 
b. {HHHHH}, {HHHHT}
c. 11
d. 5
e. 3
f. Yes, becasue they cannot both be true
g. No because they can both be true

##1.2
a. 0.2
b. 0.7
c. The expected number of young is 0.95 cats 
d. (1x0.3)+(4x0.15)+(9x0.05) = 1.35
e. Most likely it has a higher expected mean and a higher degree of variance because of the 4 value that is included.
f. 
```{r}
X <- sample(c(rep(0,1,2,3),rep(30,50,15,5)),1000000,replace=TRUE)
mean(X)
var(X)
```
g.
```{r}
Y <- sample(c(rep(0,1,2,3,4),rep(20,50,15,10,5)),1000000,replace=TRUE)
mean(Y)
var(Y)
```

##1.3
The expected value is 1 and the variance is 0 since it cannot be any other number besides 1.

##1.4
a. Yes, since they cannot both occur at the same time
b. The probability that X > 3 is 0.0 since the probability that X <= 3 is 1
c. The probability that X > 2 is greater than the probability that X > 3, because the probability that it is greater than 2 is 0.05, but the probability that it is greater than 3 is 0, acccording to the graph.

##1.5
a. 0.2
b. 0.3
c. 0, not mutually exclusive

##1.6
If a and b can occur simultaneously, then the probability that both a and b occur is P(a) + p(b)

##1.7
a. 0.5*0.5 = 0.25
b. 0.5^n
c. 1-0.5^n

##1.8
a. P(A U C)|P(B = 0)
b. P(A)|P(B U C = 0)
c. Region A since it yields the hieghest result of success

##1.9
a. Z = (X - M)/รณ = (X - 26.47413)/93.34443
b. Z(2) + z(3) = 1.85 + 2.9 = 4.758

##1.10
P(0) = 0.041
P(1) = 0.166
P(2) = 0.375
P(3) = 0.416

##1.11
a. 

##2.1
a. If the crashes are evenly distributed, I would expect each roomate to crash on average 2.6 times
```{r}
numCrashes <- 10
numRoomates <- 5
one_crash <- Vectorize(function(x){
  return(dbinom(x,numCrashes,prob = x/numCrashes))
})
one_crash(numRoomates)
```
b. 3 of the crashes seems way more likely given the distribution
c. 8 of the crashes seems horrible given the distribution, thats a whole 307 times greater than the distribution

##2.2

##2.3
```{r, eval = FALSE}
current_floor <- 9
time <- 10;
elevatorFloor <- Vectorize(function(x=1){
  x <- as.numeric(sample(1:10, 1, replace = TRUE))
  return(x)
})
elevatorTime <- Vectorize(function(x=1){
  num <- ((current_floor - elevatorFloor()) * time)
  return(num)
})

elevatorTime()
```

##2.4
a.
```{r, eval=FALSE}
experiment <- Vectorize(function(x=1){
  n01 <- 0
  n02 <- 0
  maxRuns <- 100000
  while(n01 < 40 & n02 < 40 & maxRuns > 0){
    maxRuns <- maxRuns - 1
    coin <- sample(c("H","T"),1)
    if(coin == "H"){
      n01 <- n01 + 1
      } else {
      n02 <- n02 + 1
      }
  }
  return(min(n01,n02))
})
sample.data <- experiment(1:1000)
mean(sample.data)
var(sample.data)
```
The mean will be 32.708 when the other class hits 40 and the variance is 23.23

b.
```{r, eval=FALSE}
experiment <- Vectorize(function(x=1){
  n01 <- 0
  n02 <- 0
  maxRuns <- 100000
  while(n01 < 40 & n02 < 40 & maxRuns > 0){
    maxRuns <- maxRuns - 1
    coin <- sample(c("HA","HB","TA","TB"),1, replace = FALSE)
    if(coin == "HA" & n01 > n02){
      n01 <- n01 + 1
    } else if(coin == "HA" & n02 > n01){
      n02 <- n02 + 1
    } else if(coin == "HB" & n01 > n02){
      n02 <- n02 + 1
    } else if(coin == "HB" & n02 > n01){
      n01 <- n01 + 1
    } else if(coin == "TA" & n01 > n02){
      n01 <- n01 + 1
    } else if(coin == "TA" & n02 > n01){
      n02 <- n02 + 1
    } else if(coin == "TB" & n01 > n02){
      n02 <- n02 + 1
    } else if(coin == "TB" & n02 > n01){
      n01 <- n01 + 1
    }
  }
  return(min(n01,n02))
})
sample.data <- experiment(1:10)
mean(sample.data)
var(sample.data)
```
##2.5
```{r,eval=FALSE}
N = c("Unmarked",1000, replace = FALSE)
s1 = 50
s2 = 50
experiment2 <- Vectorize(function(x=1){
  while(s1 > 1)
  {
    s1 <- s1 - 1
    N[sample(1:1000)] <- "Marked"
  }
  while(s2 > 1)
  {
    s2 <- s2 - 1
    N[sample(1:1000)] <- "Marked"
  }
  return(length(N[which("Marked")]))
})
experiment(1:100)
```

##2.6
```{r,eval=FALSE}
money = 1000
game <- Vectorize(function(x=1){
  if(money%/%3 == 0)
  {
    x <- sample(1:100)
    if(x == range(1:5))
    {
      money <- money + 1
    } else {
      money <- money - 1
    }
  }
  else
  {
    x <- sample(1:100)
    y <- sample(1:100)
    if(x == range(1:30) & y == range(1:30))
    {
      money <- money + 1
    }else if(x == range(1:30) & y != range(1:30)) {
      money <- money
    }else if(x != range(1:30) & y == range(1:30)){
      money <- money
    }else{
      money <- money - 2
    }
  }
})
game(1:1000)
```
a. In Game A, you would expect to lose money since it is not an even split of probability, in 100 attempts, you would expect to $10
b. In game B, you fluxuate around 998, but ultimately lose money
c. Sine both games you lose on average, I would expect by switching between the two randomly would net a loss in money 

##2.7
a.
```{r}
tank <- Vectorize(function(x){
  y <- sample(1:x,20,replace = FALSE)
  return(max(y))
})
N <- 150

x <- seq(150,1000, by=50)
y <- tank(x)
plot(x,y)
```
b. 142
c. 148 184 232 282 343 367 440 491 538 582 646 690 737 778 837 880 888 988

## 2.8
```{r}
darts <- c(100,1000,10000,100000)
X <- seq(0,3,length.out = 0.2)
f <- Vectorize(function(x){
  return(exp(-x^2))
})
Y <- f(X)
plot(X,Y,type='l',lwd=4)
XX <- c(X,rev(X))
YY <- c(Y,rep(0,length(X)))
polygon(XX,YY,col='red')

xmin <- 0
xmax <- 1
ymin <- 0
ymax <- 0.3
number_of_darts <- darts
X <- runif(number_of_darts,min=xmin,max=xmax)
Y <- runif(number_of_darts,min=ymin,max=ymax)
image_area <- (xmax-xmin)*(ymax-ymin)
image_area

darts_in_region <- length(which(Y <= f(X)))
darts_in_region/number_of_darts
darts_in_region/number_of_darts * image_area
```

##2.9

##2.10
```{r,eval=FALSE}
df <- data("morley")
X <- morley$Speed
mu <- mean(X)
sigma <- sd(X)

x <- seq(from=69,to=100,by=0.5)
y <- dnorm(x,mean=mu,sd=sigma)
sample_size <- length(X)

one_mean <- Vectorize(function(x=1) {
  return(mean(sample(X,sample_size,replace=TRUE)))
})
sample_means <- one_mean(1:sample_size)
meanmu <- mean(sample_means)
meansigma <- sd(sample_means)
xmean <- seq(from=78,to=83,by=0.01)
ymean <- dnorm(xmean,mean=meanmu,sd=meansigma)

interval90 <- qnorm(c(.05,.95),mean=meanmu,sd=meansigma)
interval90
```
##2.11
```{r,eval=FALSE}
df <- data("LakeHuron")
X <- LakeHuron
mu <- mean(X)
sigma <- sd(X)

x <- seq(from=69,to=100,by=0.5)
y <- dnorm(x,mean=mu,sd=sigma)
sample_size <- length(X)

one_mean <- Vectorize(function(x=1) {
  return(mean(sample(X,sample_size,replace=TRUE)))
})
sample_means <- one_mean(1:sample_size)
meanmu <- mean(sample_means)
meansigma <- sd(sample_means)
xmean <- seq(from=78,to=83,by=0.01)
ymean <- dnorm(xmean,mean=meanmu,sd=meansigma)

interval90 <- qnorm(c(.05,.95),mean=meanmu,sd=meansigma)
interval90

```

##3.1
```{r}
data("quakes")
Y <- quakes$stations
X <- sample(quakes$stations,100)
barplot(Y)
barplot(X)
```
They seem very similar, which leads me to believe that the sampling is random enough to get a fair distrbution.

##3.2
```{r}
X <- seq(from=-3,to=3,by=0.01)
Y <- pnorm(X)
plot(X,Y)
y <- runif(1)
set.seed(42)
plot(X,Y)
abline(h=y,col='red')
location <- which.min(abs(Y-y))
the_draw <- X[location]
plot(X,Y)
abline(h=y,col='red')
abline(v=the_draw,col='blue')
set.seed(42)
y <- runif(1000)
find_X <- function(y) {
location <- which.min(abs(Y-y))
  return(X[location])
}
find_X <- Vectorize(find_X)
draws <- find_X(y)
hist(draws,breaks=40)
```
