---
title: "Project1"
author: "Liam McCallum"
date: "4/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(tree)
library(class)
library(lubridate)
library(randomForest)
```

## R Markdown

```{r}
mydata <- read.csv("/Users/McCallum_Liam/Desktop/Schoolwork/Math/Math331/Project1/NyboDataSet.csv", 1)
summary(mydata)
mydata$TimeMil <- NULL
fly <- mydata[mydata$Event == "200Fly",]
IM <- mydata[mydata$Event == "200IM",]
Season1 <- subset(mydata, as.Date(mydata$Date) < "2018/3/24")
Season2 <- subset(mydata, as.Date(mydata$Date) > "2018/3/24")
mydata$Season <- ifelse(as.Date(mydata$Date) < "2018/3/24", "Season 1", "Season 2")

ggplot(mydata,aes(x = as.Date(mydata$Date),mydata$TimeSec)) +
  geom_jitter(aes(color=mydata$Season)) +
  ggtitle("Times By Season") +
  ylab("Times in Sec") +
  xlab("Date") +
  theme(legend.title = element_blank())
  
t.test(Season1$TimeSec,Season2$TimeSec)

mydata$numericDate <- as.numeric(as.Date(mydata$Date))
tree.data <- tree(TimeSec ~ numericDate,data = mydata)

plot(tree.data)
text(tree.data, pretty=0)

yhat <- predict(tree.data)
plot(yhat,mydata$TimeSec)
abline(a=0,b=1,col="red")

MSE <- function(predicted,actual) {
  sum((predicted-actual)^2)
}
MSE(yhat,mydata$TimeSec)

forestData <- randomForest(TimeSec ~ numericDate,data = mydata)

yhat2 <- predict(forestData)
plot(yhat2,mydata$TimeSec)
abline(a=0,b=1,col="red")
MSE(yhat2,mydata$TimeSec)
```

Predictive model for times for next year in both events

Let's try a linear model
```{r}
summary(lm(TimeSec~Event,data=mydata))
```
This doesn't look terrible. But some improvements could be made. Let's try a tree

```{r}
mymodel <- tree(Date ~TimeSec+Event,data=mydata)
summary(mymodel)
plot(mymodel)
text(mymodel,pretty=0)
```
This is nice but since there are not any indiciations of times per event, let's try that instead.

```{r}
mymodel <- tree(Date ~ TimeSec, data=fly)
summary(mymodel)
plot(mymodel)
text(mymodel,pretty=0)

mymodel <- tree(Date ~ TimeSec, data=IM)
summary(mymodel)
plot(mymodel)
text(mymodel,pretty=0)

```
These are very nice, even if there aren't too many terminal nodes

Let's try a random forest

```{r}
mymodel <- randomForest(Date ~ TimeSec + Event, data=mydata,mtry=2)
varImpPlot(mymodel)
```
This looks pretty nice. It shows us which variable is more important. Regardless of event, the time is a better variable. We may be able to predict it too.
```{r}
yhat <- predict(mymodel)
table(yhat)
```

This looks very nice. Age the dates and we have ourselves some nice dates our subject will swim his events. What if we do the times now.

```{r}
mymodel <- randomForest(TimeSec ~ Date+Event, data=mydata,mtry=2)
varImpPlot(mymodel)
#we can tell the Date plays a big factor, let's predict now.
yhat <- predict(mymodel)
ggplot(mydata) + 
  geom_point(aes(x=yhat,y=TimeSec,color=Season)) +
  scale_y_reverse() + scale_x_reverse() + 
  xlab("Predicted Times") + 
  ylab("Actual Times") + 
  ggtitle("Predicted Times for AJ's events")
```
We have our predictions. Now we will have to see if he has similar times next year.

Our second question:
Which event did he improve the most over the last two seasons?
```{r}
summary(lm(TimeSec~Date,data=fly))
summary(lm(TimeSec~Date,data=IM))
```
In our two linear models, it's very close, but it looks like IM had the better outcome compared to fly. The IM R-squared is .9989 compared to fly which is .9986.


